Causal inference and co-evolutionary contingencies using dynamic phylogenetic models

### 2022-06-27 manuscript.qmd
Outline popular methods

Trait co-evolution simply describes *how* traits change over time. This has value for understanding natural history and can stimulate new research questions. However, simply inferring that traits changed together in the past does not tell us *why* they changed, or how the traits would respond to intervention, i.e., causality. Nonetheless, researchers in ecology and evolution are often motivated to study trait co-evolution because they believe these patterns reflect causal relations.

>>>>

Trait co-evolution simply describes *how* traits change over time. Such descriptions have value for understanding natural history, ancestral state reconstruction (using the present to retrodict the past), and can stimulate new research questions. However, simply inferring that traits changed together in the past does not tell us *why* they changed, or how the traits would respond to intervention, i.e., causality. Nonetheless, researchers in ecology and evolution are often motivated to study trait co-evolution because they believe these patterns reflect causal relations. We briefly review some popular classes of PCM and their limitations with respect to causal inference.

## "Static" PCMs

Independent contrasts, PGLS, PGLMMs, PPA

## "Dynamic" PCMs

Pagel's discrete, Hidden Markov Models, Butler & King OU
